---
title: "R Notebook"
output: html_notebook
---
```{r}
library(lubridate)
library(tidyverse)
```

``` {R}
df_business <- readRDS("C:/Users/lilya/Downloads/R/df_business_vancouver.rds")
```

**Step 1**

Information on open hours of businesses on different days of the week are in the columns starting with hours_. For example, hours_Monday contains the open hours on Monday. These columns are of type character (i.e., strings).
In the code below, I have extracted the first 10 non-missing values of hours_Monday as the vector x. The time window that restaurants are open is specified with a particular format: h1:m1-h2:m2, with h1 and m1 denoting the hour and minute for which the restaurant opens, respectively, and h2 and m2 denoting the hour and minute for which they close, respectively.

```{r}
x <- df_business$hours_Monday
x <- x[!is.na(x)]
x <- x[1:10]
x
```

Using the vector x I retrieved the four time components (h1, m1, h2, m2) and then calculated the duration (of open time) in minutes:

```{r}
h1 <- strsplit(x, "-") %>% sapply("[", 1)
h1 <- strsplit(h1, ":") %>% sapply("[",1) %>% as.numeric()
m1 <- strsplit(x, "-") %>% sapply("[", 1)
m1 <- strsplit(m1, ":") %>% sapply("[",2) %>% as.numeric()
h2 <- strsplit(x, "-") %>% sapply("[",2)
h2 <- strsplit(h2, ":") %>% sapply("[",1) %>% as.numeric()
m2 <- strsplit(x, "-") %>% sapply("[", 2)
m2 <- strsplit(m2, ":") %>% sapply("[",2) %>% as.numeric()
duration <- (h2 * 60 + m2) - (h1 * 60 + m1) 
duration

```
**Step 2**
I made a function that takes a vector x as input—containing open hours, formatted as the hours_Monday column—and returns the duration of open time in minutes:

```{r}
get_duration <- function(x) {
  h1 <- strsplit(x, "-") %>% sapply("[", 1)
  h1 <- strsplit(h1, ":") %>% sapply("[",1) %>% as.numeric()
  m1 <- strsplit(x, "-") %>% sapply("[", 1)
  m1 <- strsplit(m1, ":") %>% sapply("[",2) %>% as.numeric()
  h2 <- strsplit(x, "-") %>% sapply("[",2)
  h2 <- strsplit(h2, ":") %>% sapply("[",1) %>% as.numeric()
  m2 <- strsplit(x, "-") %>% sapply("[", 2)
  m2 <- strsplit(m2, ":") %>% sapply("[",2) %>% as.numeric()
  out <- (h2 * 60 + m2) - (h1 * 60 + m1) 
  return(out)
}
```

**Step 3**
I used the get_duration function to calculate seven new columns for df_business, containing the duration of open time for different days of the week:

```{r}
df_duration <- df_business %>%
  mutate(
    duration_mon = get_duration(hours_Monday),
    duration_tue = get_duration(hours_Tuesday),
    duration_wed = get_duration(hours_Wednesday),
    duration_thu = get_duration(hours_Thursday),
    duration_fri = get_duration(hours_Friday),
    duration_sat = get_duration(hours_Saturday),
    duration_sun = get_duration(hours_Sunday)
    )
```

**Step 4**

First, I selected the business_id and the seven columns that contain the duration information. Then, transformed it from a wide data frame (with 8 columns) to a long data frame (with 3 columns). The 3 columns are (1) business_id, (2) wday, specifying the day of the week, and (3) duration, specifying the duration of open time for a business on a specific weekday.

```{r}
df_duration <- df_duration %>%
  select(business_id, duration_mon, duration_tue, duration_wed, duration_thu, 
         duration_fri, duration_sat, duration_sun) %>%
  pivot_longer(
    cols = -business_id,
    names_to = 'wday',
    values_to = 'duration'
    )

```

**Step 5**
Checking to see if there are any negative values in the duration column of df_duration.

```{r}
df_duration %>%
  filter(duration < 0)
```

**Step 6**
To see why we have negative duration, I take the following steps:
1. Use df_duration to select the ID of businesses with negative duration, and save the data frame as df1.
2. Use df1 to select the columns starting with hours_ of df_business.
Then, I will examine the hours_ columns of the businesses that I calculated a negative duration time for and find out the problem.

```{r}
df1 <- df_duration %>%
  filter(duration < 0) %>%
  select(business_id)
```

```{r}
df_business %>%
  inner_join(df1, by = "business_id") %>%
  select(hours_Monday, hours_Tuesday, hours_Wednesday, hours_Thursday, 
         hours_Friday, hours_Saturday, hours_Sunday)
```
**Step 7**

I wrote a new function, get_duration_2, that is similar to get_duration but also includes an if_else statement that corrects for the issue that caused negative duration. The if_else statement will change the value of h2 for the cases that we previously calculated a negative duration for.

```{R}
get_duration_2 <- function(x) {
  h1 <- strsplit(x, "-") %>% sapply("[", 1)
  h1 <- strsplit(h1, ":") %>% sapply("[",1) %>% as.numeric()
  m1 <- strsplit(x, "-") %>% sapply("[", 1)
  m1 <- strsplit(m1, ":") %>% sapply("[",2) %>% as.numeric()
  h2 <- strsplit(x, "-") %>% sapply("[",2)
  h2 <- strsplit(h2, ":") %>% sapply("[",1) %>% as.numeric()
  m2 <- strsplit(x, "-") %>% sapply("[", 2)
  m2 <- strsplit(m2, ":") %>% sapply("[",2) %>% as.numeric()
  h2 <- if_else(h2*60+m2 < h1*60+m1, h2 + 24, h2)
  out <- (h2 * 60 + m2) - (h1 * 60 + m1)
  return(out)
}
```

**Checking to see if there are negative values or not**


```{r}
df_duration2 <- df_business %>%
  mutate(
    duration_mon = get_duration_2(hours_Monday),
    duration_tue = get_duration_2(hours_Tuesday),
    duration_wed = get_duration_2(hours_Wednesday),
    duration_thu = get_duration_2(hours_Thursday),
    duration_fri = get_duration_2(hours_Friday),
    duration_sat = get_duration_2(hours_Saturday),
    duration_sun = get_duration_2(hours_Sunday)
    )
```


```{r}
df_duration2 <- df_duration2 %>%
  select(business_id, duration_mon, duration_tue, duration_wed, duration_thu, 
         duration_fri, duration_sat, duration_sun) %>%
  pivot_longer(
    cols = -business_id,
    names_to = 'wday',
    values_to = 'duration'
    )

```


```{r}
df_duration2 %>%
  filter(duration < 0)
```


**Step 8**

I calculated a data frame containing duration times called df_duration_2 similar to before, but using get_duration_2 for calculating the duration.


```{r}
df_duration_2 <- df_business %>%
  mutate(
    duration_mon = get_duration_2(hours_Monday),
    duration_tue = get_duration_2(hours_Tuesday),
    duration_wed = get_duration_2(hours_Wednesday),
    duration_thu = get_duration_2(hours_Thursday),
    duration_fri = get_duration_2(hours_Friday),
    duration_sat = get_duration_2(hours_Saturday),
    duration_sun = get_duration_2(hours_Sunday)
    )
```

```{r}
df_duration_2 <- df_duration_2 %>%
  select(business_id, duration_mon, duration_tue, duration_wed, duration_thu, 
       duration_fri, duration_sat, duration_sun) %>%
  pivot_longer(
    cols = -business_id,
    names_to = 'wday',
    values_to = 'duration'
    )
```

**Step 9** 
I change the wday column such that prefix duration_ is removed. Then, I changed the wday column into an ordinal column (type factor), with levels ordered from “mon” to “sun”. Finally, I changed duration so that it is in units of hours.

```{r}
df_duration_2 <- df_duration_2 %>%
mutate(
wday = str_replace(wday, "duration_", ""),
wday = factor(wday, order = TRUE,
              levels = c("mon", "tue", "wed", "thu", "fri", "sat", "sun")),
duration = duration %/% 60
)

```

**Step 10**
Creating the figure

```{r}
df_duration_2 %>%
  ggplot() +
  geom_hline(yintercept = 0.10) +
  geom_density(
    aes(x = duration),
    fill = "darkgoldenrod"
  ) +
  facet_wrap(~wday, ncol = 2) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 24, by = 2)) +
  labs(
  title = "Duration of time that restaurants are open",
  subtitle = "A horizontal line at y=0.10 has been added to facilitate 
  comparison",
  x = "Duration (hr)",
  y = "Density",
  caption = str_c(
    "More proportion of restaurants are closed on Mon than other days of the week",
    "\n",
    "Similar pattern on Tue, Wed, Thu, and Fri, with majority open for 8-10 hr",
    "\n",
    "Sat is similar to Mon but with less restaurants closed and more restaurants opened for < 8 hr",
    "\n",
    "A wide plateau on Sun, from 5 hr to 12 hr",
    sep = ""
    )
  ) + 
  theme(plot.caption = element_text(hjust = 0))

```

***Figure 2***

**Step 1**
I wrote a function that takes a character vector of open hours as its input and outputs closing hours as a vector. If closing hour is 2 am next day, for example, then the output should be 26.

```{r}
get_h <- function(x) {
  h1 <- strsplit(x, "-") %>% sapply("[", 1)
  h1 <- strsplit(h1, ":") %>% sapply("[",1) %>% as.numeric()
  h2 <- strsplit(x, "-") %>% sapply("[",2)
  h2 <- strsplit(h2, ":") %>% sapply("[",1) %>% as.numeric()
  h2 <- if_else(
    h2 < h1, 
    h2 + 24, 
    h2
    )
  return(h2)
}
```

**Step 2**
I made a data frame, named df_closing, that contains three columns: (1) business_id, (2) wday, and (3) closing, which specifies the closing hour. I made the column wday of type ordinal (i.e., factor) with levels from “mon” to “sun”.


```{r}
df_closing <- df_business %>%
  mutate(
    closing_mon = get_h(hours_Monday),
    closing_tue = get_h(hours_Tuesday),
    closing_wed = get_h(hours_Wednesday),
    closing_thu = get_h(hours_Thursday),
    closing_fri = get_h(hours_Friday),
    closing_sat = get_h(hours_Saturday),
    closing_sun = get_h(hours_Sunday)
    )
```

```{r}
df_closing <- df_closing %>%
  select(business_id, closing_mon, closing_tue, closing_wed, closing_thu, 
       closing_fri, closing_sat, closing_sun) %>%
  pivot_longer(
    cols = -business_id,
    names_to = 'wday',
    values_to = 'closing')
```


```{r}
df_closing <- df_closing %>%
  mutate(
  wday = str_replace(wday, "closing_", ""),
  wday = factor(wday, order = TRUE,
                levels = c("mon", "tue", "wed",
                           "thu", "fri", "sat",
                           "sun")))
```

**Step 3**
The attributes_Alcohol column of df_business has 7 unique values, but some are duplicates. For example, the values "u'none", "'none'", and "None" all specify businesses that do not serve alcohol.

```{r}
df_business$attributes_Alcohol %>% unique()
```

First, I selected the business_id and attributes_Alcohol column of df_business. Then, changed the attributes_Alcohol so that it has only three values: (1) "none", (2) "full_bar", and (3) "beer_and_wine". I saved the data frame as df_alco.

```{r}
df_alco <- df_business %>% 
  select(business_id, attributes_Alcohol) %>% 
  mutate(
    attributes_Alcohol = attributes_Alcohol %>% 
                          str_replace("u'", "'") %>%
                          str_replace_all("'","") %>%
                          tolower()
    )
```

```{r}
df_alco$attributes_Alcohol %>% unique()
```

**Step 4**
For each business in df_closing I included the information on whether it has a full bar, serves beer and wine, or does not serve any alcohol as a new column.

```{r}
df_closing <- df_closing %>%
              inner_join(df_alco, by = "business_id")
```

**Step 5**
Creating the figure

```{r}
df_closing %>%
  filter(attributes_Alcohol == "full_bar" | attributes_Alcohol == "none") %>%
  filter(closing != 0) %>%
  ggplot() +
  geom_density(
    aes(x = closing),
    fill = "darkgoldenrod"
  ) +
  facet_grid(
    wday ~ attributes_Alcohol
    ) +
  theme_bw() +
  scale_x_continuous(
    breaks = seq(0, 48, by = 2),
    labels = c(
      seq(0, 24, by = 2),
      str_c(seq(2, 24, by = 2), "\n(+1)")
      )
  ) +
  coord_cartesian(xlim = c(12, 24 + 6)) +
  labs(
    title = "Closing hours of restaurants",
    subtitle = "Restaurants with a full bar vs. those that don't serve alcohol",
    x = "Closing hour (+1 indicates next day)",
    y = "Density"
  )
```

